---
title: "Random allocation list specification"
date: "`r Sys.time()`"
output: pdf_document
params:
  randlist: NA
  blocksize: NA
  blockno: NA 
  seed: NA 
  alloclist: NA 
  stratalist: NA 
  nsites: NA
  blocks: NA
---
```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

This is the specifications of the randomisation list "`r paste0("RandList ", Sys.Date()," seed ", params$seed, ".xslx")`" produced by the https://icostatistics.shinyapp.io "Random allocation list" web application. The web application is based on R, and the code described in this document is also based on R. 

# Parameters
The following parameters define the randomisation list

## Treatment arms

There are `r nrow(params$alloclist)` treatment arms randomly allocated. The treatment arms parameter is defined as:
```{r }
params$alloclist
```
 
## Block information

`r if(params$blocks == "fixed") {paste0("The block size is fixed at ",parms$blocksize, " for all blocks.")}`

`r if(params$blocks == "simple") {"This is a simple, unblocked random allocation list. Therefore the 'block' spans the whole list."}`

`r if(params$blocks == "random") {"The block size is randomsly selected  for each block."}`


The block size and number of blocks parameters are given by:

```{r}
params$blocksize

params$blockno
```

## Number of sites

The number of sites is `r params$nsites`. The parameter is given by:

```{r}
params$nsites
```

## Stratification variables

`r if(is.null(ncol(params$stratalist))) {"There are no stratification variables."}`
`r if(!is.null(ncol(params$stratalist))) {paste0("There are ", ncol(params$stratalist), " stratification variables")}`
The stratification variable(s) parameter is given by:

```{r}
params$stratalist
```

## Seed

The seed is `r params$seed`. The parameter is given by:

```{r}
params$seed
```


# Random allocation list code

With the above parameters and the code below, the same random allocation list as produced by the application should be produced. 

```{r}
set.seed(params$seed)
arms <- nrow(params$alloclist)

my_randlist <- params$stratalist %>%
  crossing(SiteCode = 1:params$nsites) %>%
  arrange(SiteCode) %>%
  mutate(tempstrat = 1:n()) %>%
  crossing(blockno = 1:params$blockno) %>%
  mutate(blockno = 1:n()) %>%
  group_by(blockno) %>%
  mutate(
    blocksize = ifelse(
      length(params$blocksize) > 1,
      sample(params$blocksize, 1),
      params$blocksize
    ),
    blocksize = as.integer(blocksize)
  ) %>%
  mutate(blocksize2 = blocksize / as.integer(arms)) %>%
  crossing(params$alloclist) %>%
  ungroup() %>%
  select(SiteCode, blockno, blocksize, everything()) %>%
  arrange(SiteCode, blockno) %>%
  uncount(blocksize2) %>%
  group_by(blockno) %>%
  mutate(rand = runif(n())) %>%
  arrange(blockno, rand) %>%
  mutate(seq_in_block = 1:n()) %>%
  group_by(SiteCode) %>%
  mutate(RandNo = as.integer(SiteCode * 1000 + 1:n())) %>%
  select(-tempstrat,-rand) %>%
  select(SiteCode, RandNo, everything())
```

A sample of the list is as follows:

```{r}
my_randlist 
```

If everything has gone according to plan, the resulting list produced by the code specified above should produce an identical list as the one produced by the web application. A test of this is done by the following code, where the list produced by the code specified in this document is compared to the list produced by the code behind the web application. They should be identical, and the resulting test should be "TRUE".

```{r}  
vars <- c("blockno", "blocksize", "seq_in_block")
all_equal(my_randlist %>% select(-any_of(vars)) ,params$randlist%>% select(-any_of(vars)))
```

